### 2020-04-08 plosOne/S1Appendix.Rmd
figure legend

---
title: "Supplimentary"
author: "JK"
date: '2020 4 3 '
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

Data repository

Pancancer TCGA RNA data: http://api.gdc.cancer.gov/data/3586c0da-64d0-4b74-a449-5ff4d9136611

```{r TCGA abbreviations}
library(tidyverse)

read_delim("../TCGAAbbreviations.txt", delim = "\t") %>%
  knitr::kable()

```