### 2020-05-20 plosOne/S1Appendix.Rmd
author edit

---
title: "Supplimentary"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r TCGA abbreviations}
library(tidyverse)

read_delim("../TCGAAbbreviations.txt", delim = "\t") %>%
  knitr::kable()

```