### 2020-06-11 plosOne/plosOne.tex

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended
% to minimize problems and delays during our production
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% Once your paper is accepted for publication,
% PLEASE REMOVE ALL TRACKED CHANGES in this file
% and leave only the final text of your manuscript.
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file.
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment.
%
% Please add line breaks to long display equations when possible in order to fit size of the column.
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
% \usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace}
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
% \bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

\usepackage{forarray}
\usepackage{xstring}
\newcommand{\getIndex}[2]{
  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
}

\setcounter{secnumdepth}{0}

\newcommand{\getAff}[1]{
  \getIndex{#1}{Department of Hospital Pathology, Seoul St.~Mary's Hospital}
}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Prediction of PIK3CA mutation with gene expression in cancer} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Jun Kang\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital}},
Ahwon Lee\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital}},
Youn Soo Lee\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital}}\textsuperscript{*}\\
\bigskip
\textbf{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital}}Department of Hospital Pathology, Seoul St.~Mary's Hospital, College of
Medicine, The Catholic University of Korea, Seoul, South Korea\\
\bigskip
* Corresponding author: lys9908@catholic.ac.kr\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Breast cancers with PIK3CA mutations can be treated with PIK3CA
inhibitors in hormone receptor-positive HER2 negative subtypes. We
applied a supervised elastic net penalized logistic regression model to
predict PIK3CA mutations from gene expression data. This regression
approach was applied to predict modeling using the TCGA pan-cancer
dataset. Approximately 10,845 cases were available for PIK3CA mutation
and mRNA expression data. In 10-fold cross-validation, the model with
\(\lambda\) = 0.01 and \(\alpha\) = 1.0 (Ridge regression) showed the
best performance, in terms of area under the receiver operating
characteristic (AUROC). The final model was developed with selected
hyper-parameters using the entire train-set. The train-set AUROC was
0.93, and the test-set AUROC was 0.84. The area under the
precision-recall (AUPR) of the train-set was 0.66, and the test-set AUPR
was 0.39. Cancer types were the most important predictors. Both IGFR1R
and the phosphatase and tensin homolog (PTEN) were the most significant
genes in gene expression predictors. Our study suggests that predicting
genomic alterations using gene expression data is possible, with good
outcomes.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Targeted therapy has become a standard treatment for many cancer
patients, however the approach requires a test for a specific cancer
genomic alteration, to treat patients. Several direct genomic alteration
tests have been developed and proven for their clinical utility to treat
patients. Machine learning approaches can be applied to detect genomic
alterations . Machine learning algorithms can build prediction models
from a large number of predictors, such as radiomic features {[}1{]},
pathology image {[}2{]} or gene expression data {[}3{]}. Because most
direct genomic tests are more specific and sensitive than predictive
models, machine learning approaches may have limited roles in clinical
practice, however, machine learning approaches are ideal when direct
tests are unavailable or fail. RAS pathway activation predictions have
been performed using gene expression data {[}5{]}. Authors used data
from The Cancer Genome Atlas (TCGA), with a supervised elastic net
penalized logistic regression classifier, with stochastic gradient
descent. Their model performance was 84\% with an area under the
receiver operating characteristic (AUROC) curve, and 63\% with an area
under the precision-recall (AUPR) curve. Importantly, these authors
suggested their approach could be applied to other genomic alterations.
Breast cancer expressing PIK3CA mutations can be treated using PIK3CA
inhibitors, in hormone receptor-positive HER2 negative subtypes {[}6{]}.
The PIK3CA mutation is the second most common driver mutation after
TP53, and is most frequently detected in endometrial carcinoma (45\%),
followed by breast invasive carcinoma (24\%), cervical squamous cell
carcinoma, endo-cervical adenocarcinoma (20\%) and colon adenocarcinoma
(16\%). PIK3CA encodes the p110\(\alpha\) catalytic subunit of
phosphatidylinositol 3′-kinase (PI3K). PI3K is a protein kinase that
phosphorylates phosphatidylinositol 4,5-biphosphate
(PIP\textsubscript{2}) to generate phosphatidylinositol
3,4,5-triphosphate (PIP\textsubscript{3}). The phosphatase and tensin
homolog (PTEN) converts PIP\textsubscript{2} to PIP\textsubscript{3} in
contrast to PI3K. PIP\textsubscript{3} is a second messenger that
activates protein kinase B (AKT), which is a serine/threonine-specific
protein kinase. AKT inhibits apoptosis and promotes cell proliferation
{[}7{]}.

We applied a supervised elastic net penalized logistic regression model
to predict PIK3CA mutations. We wanted to ascertain whether this
prediction model approach could be applied not only to RAS pathway
activation, but also to PIK3CA mutation predictions across several
cancer types.

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{dataset}{%
\subsection{Dataset}\label{dataset}}

We used the TCGA pan-cancer dataset. TCGA archives the following; exome
sequencing, gene expression, DNA methylation, protein expression, and
clinical data from \textgreater{} 10,000 cancer samples across 33 common
cancer types. The TCGA dataset is publically available. PIK3CA mutation
data was extracted using cgdsr rpackage {[}8{]}. Gene expression data
was downloaded from the National Cancer Institute (NCI)'s Genomic Data
Commons (GDC) website. This archives data for TCGA
(https://gdc.cancer.gov/about-data/publications/pancanatlas). Gene
expression in the TCGA pan-cancer dataset is batch-corrected with
normalization. The target variable was PIK3CA mutation status. PIK3CA
status was considered positive when the case had the following PIK3CA
variants, which were the target variants of the Therascreen PIK3CA RGQ
PCR Kit ; C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R,
H1047L, H1047R, H1047Y. This kit was approved as a companion diagnostics
test to treat with PIK3CA inhibitor by the United States Food and Drug
Administration. We split three-quarters of the dataset into the
train-set and one quarter into the test-set.

\hypertarget{modeling-process}{%
\subsection{Modeling process}\label{modeling-process}}

To narrow down potential predictors, genes with a large median absolute
deviation (\textgreater{} third-quartiles) were selected. Thirty three
cancer type dummy variables were included in predictor variables.
Yeo-Johnson transformation was performed to correct skewness. Centering
and scaling were also performed. All preprocessing was performed using
the recipe r package {[}9{]}. Penalized logistic regression was applied
to prediction modeling. Ten-fold cross-validation with targe variable
stratification was performed over the hyper-parameter grid: \(\lambda\)
\{\(10^{-5}\), \(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\),
\(10^{0}\)\}, \(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Lambda (\(\lambda\))
is a penalty scaling parameter and alpha (\(\alpha\)) is a mixing
parameter of penalty function
(\((1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1\))
{[}10{]}.

\hypertarget{accessing-model-performance}{%
\subsection{Accessing model
performance}\label{accessing-model-performance}}

Model performance was evaluated using AUROC and AUPR curve approaches.
The AUPR approach is more informative than AUROC for imbalanced datasets
{[}11{]}. The modeling process and accessing model performance were
performed with the tidymodels rpackage {[}12{]}.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{dataset-summary}{%
\subsection{Dataset summary}\label{dataset-summary}}

10,845 cases were available for both PIK3CA mutation and mRNA expression
data. 5,128 out of 20,502 genes were included in the modeling process,
after filtering for median absolute deviation, as described in the
modeling process method. The prevalence rate for PIK3CA was 0.11 in all
cases. The PIK3CA prevalence rate in each cancer type varied. The median
prevalence rate of PIK3CA for each cancer type was 0.03 (range 0--0.33)
(Figure 1).

\hypertarget{selecting-model-and-performance-estimation}{%
\subsection{Selecting model and performance
estimation}\label{selecting-model-and-performance-estimation}}

For 10-fold cross-validation, the model with \(\lambda\) = 0.01 and
\(\alpha\) = 1.0 (Ridge regression) showed the best performance in terms
of AUROC. The final model was trained with the selected hyper-parameters
with the entire train-set. The train-set AUROC was 0.93 and the test-set
AUROC was 0.84. The AUPR of the train-set was 0.66 and the test-set AUPR
was 0.39 (Figure 2A).

\hypertarget{performance-of-each-cancer-type}{%
\subsection{Performance of each cancer
type}\label{performance-of-each-cancer-type}}

Because PIK3CA mutation prevalence varied across cancer types, the
performance of each cancer type was investigated. The AUROC and AUPR
were positively correlated between the train-sets and test-sets in
cancer type sub-analysis (Figure 2B). The AUPR was high in cancer types
with high PIK3CA mutation rates such as colon, breast and uterus cancer
types. The AUROC did not correlate with PIK3CA mutation rates of each
cancer type (Figure 2C).

\hypertarget{important-predictors}{%
\subsection{Important predictors}\label{important-predictors}}

The top 30 important predictors are shown (Figure 3). The coefficient is
the parameter of the predictor which represents the effect of the
predictor on prediction . IGF1R mRNA expression was the strongest
negative predictor, and PTEN was the strongest positive predictor. Both
IGFR1R and PTEN are key players in the tyrosine kinase pathway . The
cancer types were important predictors. Some cancer types including
uterine carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA),
pancreatic adenocarcinoma (PAAD), lymphoid neoplasm diffuse large B-cell
lymphoma (DLBC) were the strongest predictors.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Our model showed good performance in predicting PIK3CA mutations in
various cancer types. Our data suggested that the supervised elastic net
penalized logistic regression model could be applied not only to the RAS
activation pathway, but also to other genomic alterations. Both the RAS
activation pathway and PIK3CA mutations are key, common cancer genomic
alterations. When mutated, they exert significant effect on gene
expression in cells. However, the supervised elastic net penalized
logistic regression model cannot be generalized or applied to other
genomic alterations. It might be a challenging prediction of genomic
alterations that are infrequent or have a weak effect on gene
expression.

Prediction modeling from the TCGA pan-cancer dataset can be limiting in
terms of data preprocessing. The gene expression data is processed by
between-sample normalization to remove batch effects. If the model has
been trained from between-sample normalization, a new sample cannot be
exactly preprocessed with normalization which was done on trainset. A
model based on gene expression from the TCGA pan-cancer dataset is
limited in terms of data preprocessing. It is necessary to develop a
preprocessing method that is independent of a dataset, to apply gene
expression data to the prediction model.

Our PIK3CA prediction model was similar to the RAS activation prediction
model in terms of AUROC (0.84). However the AUPR of our model was lower
than the RAS activation model (0.39 versus 0.63). The reason for our
lower AUPR may be explained by the low prevalence rate of PIK3CA
mutations, and an imbalanced dataset. The model for RAS activation
trained with cancer types with more than 0.05 prevalence of RAS
activation to avoid imbalance classification problem. We included all
cancer types in our modeling process. The lower prevalence rate of
target variables meant our dataset had a lower AUPR baseline. In the
sub-analysis performance of each cancer type, the cancer types with
higher PIK3CA mutation rates showed better AUPRs.

Our model included cancer types as predictors, and they were stronger
predictors than gene expression. The varying prevalence of PIC3CA
mutations across cancer types may be a reason for the strong predictive
power of cancer types. If the cancer type was wrong or could not be
determined, our model performance was poor.

Some significant gene expression predictors were closely related to PTEN
and the PI3K pathway. PTEN and IGFR1R were the strongest gene expression
predictors, which has negative and positive predictive powers. IGF1R is
a tyrosine kinase receptor that activates PI3K {[}13{]}. {]}, and PTEN
is an important regulator of PIP\textsubscript{3} by dephosphorylating
PIP\textsubscript{3}, in contrast to PI3K {[}7{]}.

Several studies have attempted to predict genomic alterations from gene
expression data {[}4,14{]}. A study investigated PIK3CA mutation
predictions using gene-expression signatures which is a sum of the
average of the logarithmic gene expression . The model showed good
performance AUROC 0.71 in an independent test set {[}4,14{]}.Another
study predicted copy number alterations with gene expression, using a
multinomial logistic regression model with least absolute shrinkage and
selection operator (LASSO) parameters {[}15{]}. The prediction of the
1p/19q codel was very good, with an AUROC of 0.997, and gene-level
predictions were good, with an AUROC of 0.75 {[}15{]}. A Hidden Markov
Model Approach for Prediction of Genomic Alterations from Gene
Expression Profiling {[}16{]}. A logistic regression model was used for
MYCN gene amplification in neuroblastoma {[}17{]}.

Our study suggested that the prediction of genomic alterations using
gene expression data was possible, with good performance. However,
improved performances are required for clinical tests, and the
standardization of generation processing of gene expression data is also
needed .

\hypertarget{figure-legends}{%
\section{Figure legends}\label{figure-legends}}

\begin{itemize}
\item
  Figure 1. Prevalence rate of PIK3CA mutations across cancer types.
  Cancer type abbreviations are explained in the S1 Appendix.
\item
  Figure 2. Summary of modeling results. (A) Left: receiver operating
  characteristic (ROC) curve. Right: precision-recall (PR) curve of
  train-set and test-set. The horizontal green line is the PIK3CA
  mutation rate (0.11) (B) Correlation between train-set and test-set of
  the area under the receiver operating characteristic curve (AUROC),
  and the area under the precision-recall curve (AUPR) among cancer
  types. The gray band is the 95\% confidence interval. Abbreviations
  are explained in the S1 Appendix. (C) Correlations between the PIK3CA
  mutation rate of the AUROC, and the AUPR.
\item
  Figure 3. Coefficient model. (A) Top 30 high mRNA coefficients. (B)
  Cancer type coefficients. Cancer types abbreviations are explained in
  the S1 Appendix.
\end{itemize}

\hypertarget{supporting-information}{%
\section{Supporting information}\label{supporting-information}}

\begin{itemize}
\tightlist
\item
  S1 Appendix.
\end{itemize}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-dercleIdentificationNonSmall2020}{}%
1. Dercle L, Fronheiser M, Lu L, Du S, Hayes W, Leung DK, et al.
Identification of NonSmall Cell Lung Cancer Sensitive to Systemic Cancer
Therapies Using Radiomics. Clin Cancer Res. American Association for
Cancer Research; 2020;26: 2151--2162.
doi:\href{https://doi.org/10.1158/1078-0432.CCR-19-2942}{10.1158/1078-0432.CCR-19-2942}

\leavevmode\hypertarget{ref-coudrayClassificationMutationPrediction2018}{}%
2. Coudray N, Ocampo PS, Sakellaropoulos T, Narula N, Snuderl M, Fenyö
D, et al. Classification and mutation prediction from nonSmall cell lung
cancer histopathology images using deep learning. Nature Medicine.
Nature Publishing Group; 2018;24: 1559--1567.
doi:\href{https://doi.org/10.1038/s41591-018-0177-5}{10.1038/s41591-018-0177-5}

\leavevmode\hypertarget{ref-wayMachineLearningDetects2018}{}%
3. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
Machine Learning Detects Pan-cancer Ras Pathway Activation in The Cancer
Genome Atlas. Cell Reports. Elsevier; 2018;23: 172--180.e3.
doi:\href{https://doi.org/10.1016/j.celrep.2018.03.046}{10.1016/j.celrep.2018.03.046}

\leavevmode\hypertarget{ref-loiPIK3CAMutationsAssociated2010}{}%
4. Loi S, Haibe-Kains B, Majjaj S, Lallemand F, Durbecq V, Larsimont D,
et al. PIK3CA mutations associated with gene signature of low mTORC1
signaling and better outcomes in estrogen receptorPositive breast
cancer. PNAS. National Academy of Sciences; 2010;107: 10208--10213.
doi:\href{https://doi.org/10.1073/pnas.0907011107}{10.1073/pnas.0907011107}

\leavevmode\hypertarget{ref-way_2018_machine_cellreports}{}%
5. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
Machine Learning Detects Pan-cancer Ras Pathway Activation in The Cancer
Genome Atlas. Cell Reports. Elsevier; 2018;23: 172--180.e3.
doi:\href{https://doi.org/10.1016/j.celrep.2018.03.046}{10.1016/j.celrep.2018.03.046}

\leavevmode\hypertarget{ref-andre_2019_alpelisib_nengljmeda}{}%
6. André F, Ciruelos E, Rubovszky G, Campone M, Loibl S, Rugo HS, et al.
Alpelisib for PIK3CA-Mutated, Hormone ReceptorPositive Advanced Breast
Cancer. New England Journal of Medicine. Massachusetts Medical Society;
2019;380: 1929--1940.
doi:\href{https://doi.org/10.1056/NEJMoa1813904}{10.1056/NEJMoa1813904}

\leavevmode\hypertarget{ref-cantleyPhosphoinositide3KinasePathway2002}{}%
7. Cantley LC. The Phosphoinositide 3-Kinase Pathway. Science. American
Association for the Advancement of Science; 2002;296: 1655--1657.
doi:\href{https://doi.org/10.1126/science.296.5573.1655}{10.1126/science.296.5573.1655}

\leavevmode\hypertarget{ref-jacobsenCgdsrRbasedAPI2019}{}%
8. Jacobsen A, Luna A. Cgdsr: R-based API for accessing the MSKCC cancer
genomics data server (CGDS). 2019.

\leavevmode\hypertarget{ref-kuhnRecipesPreprocessingTools2020}{}%
9. Kuhn M, Wickham H. Recipes: Preprocessing tools to create design
matrices. 2020.

\leavevmode\hypertarget{ref-friedmanRegularizationPathsGeneralized2010}{}%
10. Friedman J, Hastie T, Tibshirani R. Regularization Paths for
Generalized Linear Models via Coordinate Descent. J Stat Softw. 2010;33:
1--22.

\leavevmode\hypertarget{ref-saitoPrecisionRecallPlotMore2015}{}%
11. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
than the ROC Plot When Evaluating Binary Classifiers on Imbalanced
Datasets. PLOS ONE. Public Library of Science; 10: e0118432.
doi:\href{https://doi.org/10.1371/journal.pone.0118432}{10.1371/journal.pone.0118432}

\leavevmode\hypertarget{ref-kuhnTidymodelsEasilyInstall2020}{}%
12. Kuhn M, Wickham H. Tidymodels: Easily install and load the
'tidymodels' packages. 2020.

\leavevmode\hypertarget{ref-leroithInsulinlikeGrowthFactor2003}{}%
13. LeRoith D, Roberts CT. The insulin-like growth factor system and
cancer. Cancer Letters. 2003;195: 127--137.
doi:\href{https://doi.org/10.1016/S0304-3835(03)00159-9}{10.1016/S0304-3835(03)00159-9}

\leavevmode\hypertarget{ref-cizkovaGeneExpressionProfiling2010}{}%
14. Cizkova M, Cizeron-Clairac G, Vacher S, Susini A, Andrieu C,
Lidereau R, et al. Gene Expression Profiling Reveals New Aspects of
PIK3CA Mutation in ERalpha-Positive Breast Cancer: Major Implication of
the Wnt Signaling Pathway. PLOS ONE. Public Library of Science; 5:
e15647.
doi:\href{https://doi.org/10.1371/journal.pone.0015647}{10.1371/journal.pone.0015647}

\leavevmode\hypertarget{ref-muCNAPEMachineLearning2019}{}%
15. Mu Q, Wang J. CNAPE: A Machine Learning Method for Copy Number
Alteration Prediction from Gene Expression. IEEE/ACM Transactions on
Computational Biology and Bioinformatics. 2019; 1--1.
doi:\href{https://doi.org/10.1109/TCBB.2019.2944827}{10.1109/TCBB.2019.2944827}

\leavevmode\hypertarget{ref-gengHiddenMarkovModel2008}{}%
16. Geng H, Ali HH, Chan WC. A Hidden Markov Model Approach for
Prediction of Genomic Alterations from Gene Expression Profiling. In:
Măndoiu I, Sunderraman R, Zelikovsky A, editors. Bioinformatics Research
and Applications. Berlin, Heidelberg: Springer; 2008. pp. 414--425.
doi:\href{https://doi.org/10.1007/978-3-540-79450-9_38}{10.1007/978-3-540-79450-9\_38}

\leavevmode\hypertarget{ref-heGeneSignaturesAssociated2020}{}%
17. He X, Qin C, Zhao Y, Zou L, Zhao H, Cheng C. Gene signatures
associated with genomic aberrations predict prognosis in neuroblastoma.
Cancer Communications. 2020;40: 105--118.
doi:\href{https://doi.org/10.1002/cac2.12016}{10.1002/cac2.12016}

\nolinenumbers

\end{document}