### 2020-09-09 S1.Table.Rmd

---
title: "S1.Table"
author: "Jun Kang"
date: '2020 9 3 '
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r hyperparameter}
library(tidymodels)

load(here::here("glmn_tune_pancancer.RData"))

tune::show_best(glmn_tune, metric = "roc_auc")
```