### 2020-09-09 plosOne/S1Appendix.Rmd

---
title: "Supplimentary"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r TCGA abbreviations}
library(tidyverse)

read_delim(here::here("TCGAAbbreviations.txt"), delim = "\t") %>%
  knitr::kable()

```

```{r select model}

library(tidymodels)
library(latex2exp)

load(here::here("glmn_tune_pancancer.RData"))

roc_vals <- 
  collect_metrics(glmn_tune) %>% 
  filter(.metric == "roc_auc")
roc_vals %>% 
  mutate(mixture = format(mixture)) %>% 
  ggplot(aes(x = penalty, y = mean, col = mixture)) + 
  geom_line() + 
  geom_point() + 
  geom_errorbar(aes(ymin = mean - std_err, ymax = mean + std_err, width = 0.05)) +
  scale_x_log10() +
  xlab(TeX("Penalty ($\\lambda$)")) +
  ylab("Area under the receiver operating characteristic (AUROC)") +
  labs(color = TeX("Mixture ($\\alpha$)"))

```